---
title: "Results"
author: "Holloway, O., Bindoff, A., Ziebell, J."
output: html_document
---

`r Sys.Date()`  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)
```


```{r}
library(readr)
library(dplyr)
library(lme4)
library(car)
library(ggplot2)

set.seed(123)
nsim <- 100

df <- read_csv("MicroglialMorphology.csv")
#View(df)
df <- na.omit(df)
df$agemonth <- factor(df$agemonth)
df$strain <- factor(df$strain, levels = c("wt", "tg"))

xtabs(~agemonth + strain, df)
xtabs(~plaque + strain + agemonth, df)
xtabs(~location + strain, df)
```

### Statistical Analysis

To account for clustering at animal level, mixed effects logistic regression was used to model the proportion of cells with each morphology per section. Type III Wald chi-square tests were used to determine statistical significance at $\alpha = .05$. All statistical analysis was computed using the R statistical language (R Core Team, 2016; v3.3.1).

References:
R Core Team (2016). R: A language and environment for statistical computing. R Foundation for Statistical
  Computing, Vienna, Austria. URL https://www.R-project.org/

Hypothesis 1:  increased dystrophic / degenerating microglia (hyper-ramified and amoeboid morphologies) associated with AD disease progression.


```{r}
df0 <- filter(df, strain == "tg")
m1 <- glmer(cbind(ramified, totalcells) ~ agemonth + (1|animal),
                data = df0, family = "binomial")
Anova(m1, type = 3)
aggregate(ramified/totalcells ~ agemonth, df0, mean)

ggplot(df0, aes(x = agemonth, y = ramified/totalcells*100, colour = location)) +
  geom_dotplot(binaxis='y', stackdir='center',
                position = "dodge",
               aes(fill = location),
               dotsize = 3/5, alpha = 3/5) +
   theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
labs(y = "% Ramified / section", x = "Age (months)")
```



```{r}
m2 <- glmer(cbind(hyperamified, totalcells) ~ agemonth + (1|animal),
                data = df0, family = "binomial")
Anova(m2, type = 3)
aggregate(hyperamified/totalcells ~ agemonth, df0, mean)

ggplot(df0, aes(x = agemonth, y = hyperamified/totalcells*100, colour = location)) +
  geom_dotplot(binaxis='y', stackdir='center',
                position = "dodge",
               aes(fill = location),
               dotsize = 3/5, alpha = 3/5) +
   theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
labs(y = "% Hyperamified / section", x = "Age (months)")
```

```{r}
# m2.fe <- glm(cbind(hyperamified, totalcells) ~ location + agemonth,
#                 data = df0, family = "binomial")
# df.hr <- df0 
# l.hat <- predict(m2.fe, df.hr, se.fit = T)
# c.lwr <- c(l.hat$fit-1.96*l.hat$se.fit)
# c.upr <- c(l.hat$fit+1.96*l.hat$se.fit)
# df.hr$lwr <- exp(c.lwr)/(1+exp(c.lwr))
# df.hr$upr <- exp(c.upr)/(1+exp(c.upr))
# df.hr$phyperamified <- df.hr$hyperamified/df.hr$totalcells
# 
# 
# 
# ggplot(df.hr, aes(x = agemonth, y = phyperamified, colour = location)) +
#   geom_dotplot(binaxis='y', stackdir='center',
#                 position = "dodge",
#                aes(fill = location),
#                dotsize = 3/5, alpha = 1/2) +
#   geom_errorbar(position = "dodge",
#                   aes(ymax=upr, ymin=lwr), width=.5) + 
#   theme(panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.background = element_blank(), 
#         axis.line = element_line(colour = "black")) 










m3 <- glmer(cbind(amoeboid, totalcells) ~ agemonth + (1|animal),
                data = df0, family = "binomial")
Anova(m3, type = 3)
aggregate(amoeboid/totalcells ~ agemonth, df0, mean)

ggplot(df0, aes(x = agemonth, y = amoeboid/totalcells*100, colour = location)) +
  geom_dotplot(binaxis='y', stackdir='center',
                position = "dodge",
               aes(fill = location),
               dotsize = 3/5, alpha = 3/5) +
   theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
labs(y = "% Amoeboid / section", x = "Age (months)")

```

Hypothesis 2:  increase in dystrophic microglia in aging WT samples but to a lesser extent than the AD model.

```{r}
df0 <- df
m1 <- glmer(cbind(ramified, totalcells) ~ agemonth*strain + (1|animal),
                data = df0, family = "binomial")
Anova(m1, type = 3)
aggregate(ramified/totalcells ~ agemonth*strain, df0, mean)
aggregate(ramified/totalcells ~ agemonth, df0, mean)
aggregate(ramified/totalcells ~ strain, df0, mean)



ggplot(df0, aes(x = agemonth, y = ramified/totalcells*100, colour = strain)) +
  geom_dotplot(binaxis='y', stackdir='center',
                position = "dodge",
               aes(fill = strain),
               dotsize = 3/5, alpha = 3/5) +
   theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
labs(y = "% Ramified / section", x = "Age (months)")
```


```{r}
m2 <- glmer(cbind(hyperamified, totalcells) ~ agemonth*strain + (1|animal),
                data = df0, family = "binomial")
Anova(m2, type = 3)
aggregate(hyperamified/totalcells ~ agemonth*strain, df0, mean)
aggregate(hyperamified/totalcells ~ agemonth, df0, mean)
aggregate(hyperamified/totalcells ~ strain, df0, mean)

ggplot(df0, aes(x = agemonth, y = hyperamified/totalcells*100, colour = strain)) +
  geom_dotplot(binaxis='y', stackdir='center',
                position = "dodge",
               aes(fill = strain),
               dotsize = 3/5, alpha = 3/5) +
   theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
labs(y = "% Hyperamified / section", x = "Age (months)")



m3 <- glmer(cbind(amoeboid, totalcells) ~ agemonth*strain + (1|animal),
                data = df0, family = "binomial")
Anova(m3, type = 3)
m4 <- glmer(cbind(amoeboid, totalcells) ~ agemonth + strain + (1|animal),
            data = df0, family = "binomial")
anova(m3, m4)
```


A likelihood-ratio test did not support inclusion of the interaction $(\chi^2_2 = 2.73, p = .26)$, so a test of the main effects was conducted. This showed a significant main effect of age $(\chi^2_2 = 17.5, p < .001)$ and strain $(\chi^2_1 = 19.9, p < .001)$, with the proportion of amoeboid cells per section increasing from 0.5% at 3 months to 3.9% at 6 months, and 10.7% at 12 months. Amoeboid cells made up 9.6% of cells in sections from tg mice, but only .5% of cells in sections from wt mice.


```{r}
Anova(m4)
aggregate(amoeboid/totalcells ~ agemonth, df0, mean)
aggregate(amoeboid/totalcells ~ strain, df0, mean)
aggregate(amoeboid/totalcells ~ agemonth + strain, df0, mean)


ggplot(df0, aes(x = agemonth, y = amoeboid/totalcells*100, colour = strain)) +
  geom_dotplot(binaxis='y', stackdir='center',
                position = "dodge",
               aes(fill = strain),
               dotsize = 3/5, alpha = 3/5) +
   theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
labs(y = "% Amoeboid / section", x = "Age (months)")

```

Hypothesis 3:  higher proportion of dystrophic microglia surrounding plaques

```{r}
df0 <- filter(df, strain == "tg")
m1 <- glmer(cbind(ramified, totalcells) ~ location + (1|animal),
                data = df, family = "binomial")
Anova(m1, type = 3)
aggregate(ramified/totalcells ~ location, df0, mean)

m2 <- glmer(cbind(hyperamified, totalcells) ~ location + (1|animal),
                data = df, family = "binomial")
Anova(m2, type = 3)
aggregate(hyperamified/totalcells ~ location, df0, mean)

m3 <- glmer(cbind(amoeboid, totalcells) ~ location + (1|animal),
                data = df, family = "binomial")
Anova(m3, type = 3)
aggregate(amoeboid/totalcells ~ location, df0, mean)

```




